---
title: "Advanced Informatics Week 3 lab"
output: html_document
---

## dplyr

```{r setup, include=FALSE}
library(reticulate)
library(printr)
use_python("/Users/liz/miniconda3/envs/new_env/bin/python",required = T)
knitr::opts_chunk$set(echo = TRUE)
```

### Working with small dataset by loading into memory
```{r results='asis'}
data(mtcars)
help(mtcars)
```

The first several rows are:
```{r}
head(mtcars)
```

Group mtcars by `cyl` and get the mean `mpg` (first example):
```{r}
aggregate(mpg ~ cyl, data=mtcars, mean)
```

Get mean `mpg` for all `(cyl, gear)` combinations (second example):
```{r}
aggregate(mpg ~ cyl + gear, data=mtcars, mean)
```

The `dplyr` version of the first example:
```{r collapse = TRUE}
library(dplyr)
results = mtcars %>%
    group_by(cyl) %>%
    summarise(mean_mpg = mean(mpg))
results
```

The `dplyr` version of the second example:
```{r collapse = TRUE}
results = mtcars %>%
    group_by(cyl, gear) %>%
    summarise(mean_mpg = mean(mpg))
as.data.frame(results)
```

### Working with the same dataset in a relational database

Get data into an `sqlite3` database:
```{r collapse = TRUE}
library(dbplyr)
# Create a connection ("con") to a database file:
con <- DBI::dbConnect(RSQLite::SQLite(), "/Users/liz/Documents/AdvancedInformatics_RMarkdownWeek3/mtcars.sqlite3")
# Write our data frame to the database in a table called "mtcars"
DBI::dbWriteTable(con, "mtcars", mtcars, overwrite = T)
# Disconnect from our database
DBI::dbDisconnect(con)
```

```{bash, engine.opts='-l', collapse=TRUE}
ls -lhrt /Users/liz/Documents/AdvancedInformatics_RMarkdown/*.sqlite3
```

##### Analyze database using dplyr
```{r}
con <- DBI::dbConnect(RSQLite::SQLite(), "mtcars.sqlite3")
mtcars2 <- tbl(con, "mtcars")
g = mtcars2 %>% 
    group_by(cyl) %>%
    summarise(mean_mpg=mean(mpg))
```

```{r, collapse=TRUE}
g %>% show_query()
```

```{r}
result = g %>% collect()
as.data.frame(result)
```

## Python 

### Working with small dataset by loading into memory

Get `mtcars` into Python:
```{python, collapse=TRUE}
mtcars = r.mtcars
print(mtcars.head())
```

```{r, echo=F}
if (file.exists("/Users/liz/Documents/AdvancedInformatics_RMarkdownWeek3/mtcars_from_pandas.sqlite3"))
{
    file.remove("/Users/liz/Documents/AdvancedInformatics_RMarkdownWeek3/mtcars_from_pandas.sqlite3")
}
```

Repeat split/apply/combine analysis example 1 in Python:
```{python, collapse=TRUE}
mtcars.groupby(['cyl'])['mpg'].mean()
```

Repeat example 2 in Python:
```{python, collapse=TRUE}
mtcars.groupby(['cyl', 'gear'])['mpg'].mean()
```

### Working with the same dataset in a relational database

#### Analyze database with pandas
```{python}
import sqlite3 # Built into the Python language!
con = sqlite3.connect("/Users/liz/Documents/AdvancedInformatics_RMarkdownWeek3/mtcars_from_pandas.sqlite3")
# Add our data frame to the mtcars table in the database
mtcars.to_sql("mtcars", con)
con.close()
```

Checking that it worked:
```{bash, engine.opts='-l', collapse=TRUE}
ls -lhrt /Users/liz/Documents/AdvancedInformatics_RMarkdown/*.sqlite3
```

#### Reading it back in
```{python, collapse=TRUE}
import pandas as pd

con = sqlite3.connect("/Users/liz/Documents/AdvancedInformatics_RMarkdownWeek3/mtcars_from_pandas.sqlite3")
df = pd.read_sql("select * from mtcars", con)
print(df.head())
```

Have to use raw `SQL` in Python...most of the time `dplyr` is easier.
```{python, collapse=TRUE}
df = pd.read_sql("select cyl, avg(mpg) from mtcars group by cyl", con)
print(df.head())
df = pd.read_sql("select cyl, gear, avg(mpg) from mtcars group by cyl, gear", con)
print(df.head())
```

